Anomalous waves triggered by abrupt depth changes: laboratory experiments and truncated KdV statistical mechanics

### 2019-11-07 paper2_JNLS/paper.tex
NM

\begin{abstract} 
Recent laboratory experiments revealed the emergence of anomalous statistics in a randomized surface waves upon encountering an abrupt depth change (ADC) in the bottom topography (Bolles et al. PRF 2019). Majda et al.~2019 established a theoretical foundation, based on analysis of the truncated KdV (TKdV) system, that successfully captures several key qualitative features of the experiments including the robust emergence of anomalous statistics. The framework exploits the Hamiltonian structure of TKdV to construct invariant measures that are statistical matched at the ADC, and is corroborated by simulations of the deterministic dynamics. 
%
Here we present a more comprehensive treatment of these parallel experimental and modeling efforts, with emphasis on the link between the two. We present an assortment of new experimental measurements and carefully compare them to model predictions. We establish a precise link between experimental and model parameters, so that the invariant measure underlying the model can be calibrated to the specific conditions present in experiments. The model predicts a peculiar relationship between surface-displacement skewness and surface-slope variance, which is confirmed in spectacular fashion by direct experimental measurements.
\end{abstract}

@@

In this section, the variable-depth KdV equation \eqref{KdV} will be recast into a dimensionless form that is chosen for convenience in working with the statistical-mechanics framework from Majda et al.~2019 \cite{majda2019}. Since the choice of normalization not unique, it is instructive to first introduce a generic normalization to facilitate comparison with other possible choices. To this end, consider characteristic scales, $\ampscale, \lengthscale, \timescale$ for the wave amplitude, longitudinal length, and time respectively, which can remain unspecified for the moment. We introduce the dimensionless variables
\begin{align}
&u = \eta / \ampscale
&&\mbox{\em dimensionless surface displacement} \\
&\tilde{x} = (x-ct) / \lengthscale
&&\mbox{\em dimensionless position (in moving frame)} \\
&\tilde{t} = t / \timescale
&&\mbox{\em dimensionless time}
\end{align}
Recasting \eqref{KdV} in terms of these variables gives the generic dimensionless KdV equation:
\begin{equation}
u_t + \frac{3}{2} \left( \frac{c \timescale \ampscale}{\lengthscale \depth} \right) u u_x 
+ \frac{1}{6} \left( \frac{c \timescale \depth^2}{\lengthscale^3} \right) u_{xxx} = 0
\end{equation}
We have dropped the tilde notation above for simplicity and will henceforth use tildes only in cases of possible ambiguity.

@@

% NOTES
% 1) The PNAS paper uses C_theta^{1}. The Majda, Qi Stat Phys paper uses C_theta sometimes and C_theta^{-1} in other places (page 11). So I'll just pick the one I like better.
% 2) Could comment that other possible ensembles exist \cite{kleeman2014nonequilibrium}.

>>>>

\begin{abstract} 
Recent laboratory experiments revealed the emergence of anomalous statistics in a randomized surface waves upon encountering an abrupt depth change (ADC) in the bottom topography (Bolles et al. PRF 2019). Majda et al.~2019 established a theoretical foundation, based on analysis of the truncated KdV (TKdV) system, that successfully captures several key qualitative features of the experiments including the robust emergence of anomalous statistics. The framework exploits the Hamiltonian structure of TKdV to construct invariant measures that are statistical matched at the ADC, and is corroborated by simulations of the deterministic dynamics. 
%
Here we present a more comprehensive treatment of these parallel experimental and modeling efforts, with emphasis on the link between the two. We present an assortment of new experimental measurements and carefully compare them to model predictions. We establish a precise link between experimental and model parameters, so that the invariant measure underlying the model can be calibrated to the specific conditions present in experiments. The model predicts a peculiar relationship between surface-displacement skewness and surface-slope variance, which is confirmed in a spectacular fashion by direct experimental measurements.
\end{abstract}

@@

In this section, the variable-depth KdV equation \eqref{KdV} will be recast into a dimensionless form that is chosen for convenience in working with the statistical-mechanics framework from Majda et al.~2019 \cite{majda2019}. Since the choice of normalization is not unique, it is instructive to first introduce a generic normalization to facilitate comparison with other possible choices. To this end, we consider characteristic scales, $\ampscale, \lengthscale, \timescale$ for the wave amplitude, longitudinal length, and time respectively, which can remain unspecified for the moment. We introduce the dimensionless variables
\begin{align}
&u = \eta / \ampscale
&&\mbox{\em dimensionless surface displacement} \\
&\tilde{x} = (x-ct) / \lengthscale
&&\mbox{\em dimensionless position (in moving frame)} \\
&\tilde{t} = t / \timescale
&&\mbox{\em dimensionless time}
\end{align}
Recasting \eqref{KdV} in terms of these variables gives the generic dimensionless KdV equation:
\begin{equation}
u_t + \frac{3}{2} \left( \frac{c \timescale \ampscale}{\lengthscale \depth} \right) u u_x 
+ \frac{1}{6} \left( \frac{c \timescale \depth^2}{\lengthscale^3} \right) u_{xxx} = 0
\end{equation}
We have dropped the tilde notation above for simplicity and will henceforth use tildes only in cases of possible ambiguity.

@@

	We note that, in the current formulation, a positive inverse temperature $\theta > 0$ produces physically realistic statistics with a decaying energy spectrum, as is consistent with experiments \cite{majda2019, bajars2013weakly}. Negative inverse temperature produces a physically unrealistic spectrum that has more energy at smaller scales. Hence, we will hereafter focus on the physically realistic case of $\theta > 0$.
